---
title: "StatisticalAnalysis"
output: html_document
---

```{r, message=F, warning=FALSE, echo=FALSE}
library(lme4) #lmer model
library(lmerTest) #p-values
library(effects) #plot effects
library(MuMIn) #r.squaredGLMM
library(merTools) #predictInterval
library(lattice) #xyplot
library(ggplot2) #xyplot

Navafria <- read.csv("Navafria.csv", sep=";")

# To calculate SP (Summer PRecipitation Index)
Navafria$SDI <- (max(Navafria$SuPrecSum)-Navafria$SuPrecSum)/(max(Navafria$SuPrecSum)-min(Navafria$SuPrecSum))

# To standardise predicted variables: fc. scale
Navafria$AltitudeS = scale(Navafria$Altitude, scale=T) # scale
Navafria$V1s = scale(Navafria$V1, scale=T)
Navafria$WiTemps = scale(Navafria$WiTemp, scale=T)
Navafria$SDIs = scale(Navafria$SDI, scale=T)

#Save as new data file 
save(Navafria, file = "Navafria.Rdata")
```


```{r}
# Transformation functions for the response variables

transformGrowth <- function(x){
  me = mean(x)
  sig = sign(x - me)
  val = (abs(x - me) +1)^0.5
  return(val * sig - sig + me)
}

transformRecruitment <- function(x){
  me = mean(x)
  sig = sign(x - me)
  val = (abs(x - me) +1)^0.2
  out = val * sig - sig+ me
  out =  sqrt(out - min(out) + 1)
  return(out)
}

# Backtransformation functions

backTransformGrowth <- function(x){
  me = mean(x)
  sig = sign(x - me)
  val = (abs(x - me) +1)^(1/0.5)
  return(val * sig - sig + me)
}

backTransformRecruitment <- function(x){
  x = (x - min(x) + 1)^(1/0.5) - 1 + min(x)
  me = mean(x)
  sig = sign(x - me)
  val = (abs(x - me) +1)^(1/0.2)
  return(val * sig - sig + me)
}

```


```{r,warning=FALSE}

# Growth model - Fixed hypothesis

FitG = lmer(transformGrowth(GrowthReal) ~ V1s*SDIs+AltitudeS*V1s+WiTemps+(1|ID), data = Navafria, na.action = "na.exclude")
summary(FitG)
plot(allEffects(FitG))
efG <- effect("V1s:SDIs", FitG) 
plot(efG, x.var = "V1s")

# Get R2 from the library piecewiseSEM (Jonathan S. Lefcheck)
r.squaredGLMM(FitG)

# Analysis of residuals
ResG <- residuals(FitG)
FittedG <- fitted(FitG)
par(mfrow=c(2,2))
plot(ResG ~ FittedG, xlab="Fitted values", ylab="Residuals", main="Residuals vs. fitted")
abline(h=0) ## Homocedasticity
plot(Navafria$GrowthReal ~ FittedG, xlab="FittedG", ylab="GrowthReal", main = "xyplot")
abline(h=0)
hist(ResG, main="Histogram of residuals", xlab="Residuals") ## Normality
qqnorm(ResG) # Normality
qqline(ResG)

xyplot(GrowthReal ~ FittedG, data = Navafria, type = c("p","r"), col.line = "red")

# Check normality with qqplot or Shapiro-Wilk test
shapiro.test(residuals(FitG))


# Recruitment model - Fixed hypothesis
FitR = lmer(transformRecruitment(RecruitmentReal) ~ V1s*SDIs+AltitudeS*V1s+WiTemps+(1|ID), data = Navafria, na.action = "na.exclude")
summary(FitR)
plot(allEffects(FitR))
efR <- effect("V1s:SDIs", FitR) 
plot(efR, x.var = "V1s")

# Get R2 from the library piecewiseSEM (Jonathan S. Lefcheck)
r.squaredGLMM(FitR)

# Analysis of residuals
ResR <- residuals(FitR)
FittedR <- fitted(FitR) 
par(mfrow=c(2,2))
plot(ResR ~ FittedR, xlab="Fitted values", ylab="Residuals", main="Residuals vs. fitted")
abline(h=0) ## Homocedasticity
plot(Navafria$RecruitmentReal ~ FittedR, xlab="FittedR", ylab="RecruitmentReal", main = "xyplot")
abline(h=0)
hist(ResR, main="Histogram of residuals", xlab="Residuals") ## Normality
qqnorm(ResR)
qqline(ResR)

xyplot(RecruitmentReal ~ FittedG, data = Navafria, type = c("p","r"), col.line = "red")

# Check normality with qqplot or Shapiro-Wilk test
shapiro.test(residuals(FitR))

```

```{r cars}

# Graph effect
efdataG <- as.data.frame(efG)
efdataG1 <- subset(efdataG, SDIs==-1)
efdataG2 <- subset(efdataG, SDIs==2)

efdataR <- as.data.frame(efR)
efdataR1 <- subset(efdataR, SDIs==-1)
efdataR2 <- subset(efdataR, SDIs==2)

par(mfrow=c(2,1),mar=c(4,5,1,5)+.1)

hist(Navafria$V1s,axes=F,ylab="",xlab="",breaks=15,col="gray93",border="gray90",main="")
axis(side = 4,at=c(0,100))
axis(side = 4)
mtext(side = 4, line = 3, 'Frequency of V1')
par(new=T)
plot(fit~V1s,data=efdataG1,type="l",ylim=c(1,5),ylab="",xlab="",xaxt="n",bty="n")
axis(1,at=c(-1:3),labels=c("0","125","250","375","500"),tck=-0.025)
par(new=T)
polygon(c(efdataG1$V1s,rev(efdataG1$V1s)),c(efdataG1$upper,rev(efdataG1$fit)),col=rgb(0,0,1,0.2),border="white")
polygon(c(efdataG1$V1s,rev(efdataG1$V1s)),c(efdataG1$lower,rev(efdataG1$fit)),col=rgb(0,0,1,0.2),border="white")
par(new=T)
plot(fit~V1s,data=efdataG1,type="l",ylim=c(1,5),ylab="trans(Growth)",
     xlab="",lwd=2,xaxt="n",bty="n")
par(new=T)
plot(fit~V1s,data=efdataG2,type="l",ylim=c(1,5),ylab="",xlab="",lwd=2,xaxt="n",bty="n")
par(new=T)
polygon(c(efdataG2$V1s,rev(efdataG2$V1s)),c(efdataG2$upper,rev(efdataG2$fit)),col=rgb(1,0,0,0.2),border="white")
par(new=T)
polygon(c(efdataG2$V1s,rev(efdataG2$V1s)),c(efdataG2$lower,rev(efdataG2$fit)),col=rgb(1,0,0,0.2),border="white")
par(new=T)
plot(fit~V1s,data=efdataG2,type="l",ylim=c(1,5),ylab="",xlab="",lwd=2,lty=2,xaxt="n",bty="n")
title("a)",adj=0,font.main=2,cex.main=1)

hist(Navafria$V1s,axes=F,ylab="",xlab="",breaks=15,col="gray93",border="gray90",main="")
axis(side = 4,at=c(0,100))
axis(side = 4)
mtext(side = 4, line = 3, 'Frequency of V1')
par(new=T)
plot(fit~V1s,data=efdataR1,type="l",ylim=c(0.8,1.8),ylab="",xlab="",xaxt="n",bty="n")
axis(1,at=c(-1:3),labels=c("0","125","250","375","500"),tck=-0.025)
par(new=T)
polygon(c(efdataR1$V1s,rev(efdataR1$V1s)),c(efdataR1$upper,rev(efdataR1$fit)),col=rgb(0,0,1,0.2),border="white")
polygon(c(efdataR1$V1s,rev(efdataR1$V1s)),c(efdataR1$lower,rev(efdataR1$fit)),col=rgb(0,0,1,0.2),border="white")
par(new=T)
plot(fit~V1s,data=efdataR1,type="l",ylim=c(0.8,1.8),ylab="trans(Ingrowth)",
     xlab=expression(paste("V1 (", m^3, "/", ha,")", sep="")),lwd=2,xaxt="n",bty="n")
par(new=T)
plot(fit~V1s,data=efdataR2,type="l",ylim=c(0.8,1.8),ylab="",xlab="",lwd=2,xaxt="n",bty="n")
par(new=T)
polygon(c(efdataR2$V1s,rev(efdataR2$V1s)),c(efdataR2$upper,rev(efdataR2$fit)),col=rgb(1,0,0,0.2),border="white")
par(new=T)
polygon(c(efdataR2$V1s,rev(efdataR2$V1s)),c(efdataR2$lower,rev(efdataR2$fit)),col=rgb(1,0,0,0.2),border="white")
par(new=T)
plot(fit~V1s,data=efdataR2,type="l",ylim=c(0.8,1.8),ylab="",xlab="",lwd=2,lty=2,xaxt="n",bty="n")
title("b)",adj=0,font.main=2,cex.main=1)

legend('bottom','groups',c("Low SDI","High SDI"), lty = c(1,2),
       col=c('black','black'),ncol=2,bty ="n")


###### Backtransform effects

# Graph effect

backefG <- apply(efdataG[,c(3:6)], 2, backTransformGrowth)
backefG <- cbind(efdataG[,c(1:2)],backefG)
backefG1 <- subset(backefG, SDIs==-1&V1s==-1|SDIs==-1&V1s==3)
backefG2 <- subset(backefG, SDIs==2&V1s==-1|SDIs==2&V1s==3)

backefR <- apply(efdataR[,c(3:6)], 2, backTransformRecruitment)
backefR <- cbind(efdataR[,c(1:2)],backefR)
backefR1 <- subset(backefR, SDIs==-1&V1s==-1|SDIs==-1&V1s==3)
backefR2 <- subset(backefR, SDIs==2&V1s==-1|SDIs==2&V1s==3)


par(mfrow=c(2,1),mar=c(4,5,1,5)+.1)

hist(Navafria$V1s,axes=F,ylab="",xlab="",breaks=15,col="gray93",border="gray90",main="")
axis(side = 4,at=c(0,100))
axis(side = 4)
mtext(side = 4, line = 3, 'Frequency of V1')
par(new=T)
plot(fit~V1s,data=backefG1,type="l",ylim=c(-6,6),ylab="",xlab="",xaxt="n",bty="n")
axis(1,at=c(-1:3),labels=c("0","125","250","375","500"),tck=-0.025)
par(new=T)
polygon(c(backefG1$V1s,rev(backefG1$V1s)),c(backefG1$upper,rev(backefG1$fit)),col=rgb(0,0,1,0.2),border="white")
polygon(c(backefG1$V1s,rev(backefG1$V1s)),c(backefG1$lower,rev(backefG1$fit)),col=rgb(0,0,1,0.2),border="white")
par(new=T)
plot(fit~V1s,data=backefG1,type="l",ylim=c(-6,6),ylab=expression(paste("Growth (", m^3, "/", ha,")", sep="")),xlab="",lwd=2,xaxt="n",bty="n")
par(new=T)
plot(fit~V1s,data=backefG2,type="l",ylim=c(-6,6),ylab="",xlab="",lwd=2,xaxt="n",bty="n")
par(new=T)
polygon(c(backefG2$V1s,rev(backefG2$V1s)),c(backefG2$upper,rev(backefG2$fit)),col=rgb(1,0,0,0.2),border="white")
par(new=T)
polygon(c(backefG2$V1s,rev(backefG2$V1s)),c(backefG2$lower,rev(backefG2$fit)),col=rgb(1,0,0,0.2),border="white")
par(new=T)
plot(fit~V1s,data=backefG2,type="l",ylim=c(-6,6),ylab="",xlab="",lwd=2,lty=2,xaxt="n",bty="n")
title("a)",adj=0,font.main=2,cex.main=1)

hist(Navafria$V1s,axes=F,ylab="",xlab="",breaks=15,col="gray93",border="gray90",main="")
axis(side = 4,at=c(0,100))
axis(side = 4)
mtext(side = 4, line = 3, 'Frequency of V1')
par(new=T)
plot(fit~V1s,data=backefR1,type="l",ylim=c(-15,15),ylab="",xlab="",xaxt="n",bty="n")
axis(1,at=c(-1:3),labels=c("0","125","250","375","500"),tck=-0.025)
par(new=T)
polygon(c(backefR1$V1s,rev(backefR1$V1s)),c(backefR1$upper,rev(backefR1$fit)),col=rgb(0,0,1,0.2),border="white")
polygon(c(backefR1$V1s,rev(backefR1$V1s)),c(backefR1$lower,rev(backefR1$fit)),col=rgb(0,0,1,0.2),border="white")
par(new=T)
plot(fit~V1s,data=backefR1,type="l",ylim=c(-15,15),ylab=expression(paste("Ingrowth (", nºindiv , "/", ha,")", sep="")),xlab=expression(paste("V1 (", m^3, "/", ha,")", sep="")),lwd=2,xaxt="n",bty="n")
par(new=T)
plot(fit~V1s,data=backefR2,type="l",ylim=c(-15,15),ylab="",xlab="",lwd=2,xaxt="n",bty="n")
par(new=T)
polygon(c(backefR2$V1s,rev(backefR2$V1s)),c(backefR2$upper,rev(backefR2$fit)),col=rgb(1,0,0,0.2),border="white")
par(new=T)
polygon(c(backefR2$V1s,rev(backefR2$V1s)),c(backefR2$lower,rev(backefR2$fit)),col=rgb(1,0,0,0.2),border="white")
par(new=T)
plot(fit~V1s,data=backefR2,type="l",ylim=c(-15,15),ylab="",xlab="",lwd=2,lty=2,xaxt="n",bty="n")
title("b)",adj=0,font.main=2,cex.main=1)

legend('bottom','groups',c("Low SDI","High SDI"), lty = c(1,2),
       col=c('black','black'),ncol=2,bty ="n")


# Other options for graphs --> ggplot

backefGtry <- rbind(backefG1,backefG2)
row.names(backefGtry) <- 1:nrow(backefGtry)
a <- c("Low volume","High volume","Low volume","High volume")
b <- c("Low SDI","Low SDI","High SDI","High SDI")
ab <- as.data.frame(cbind(a,b))
backefGtry <- cbind(ab,backefGtry)
colnames(backefGtry)<- c("V1","SDI","V1s","SDIs","Growth","se","lower","upper")

backefRtry <- rbind(backefR1,backefR2)
row.names(backefRtry) <- 1:nrow(backefRtry)
backefRtry <- cbind(ab,backefRtry)
colnames(backefRtry)<- c("V1","SDI","V1s","SDIs","Ingroth","se","lower","upper")

ggplot(backefGtry, aes(SDI, Growth, color=V1, fill=V1)) + geom_bar(stat="identity", position="dodge") + geom_errorbar(aes(ymin=Growth-se, ymax=Growth+se), width=0.4, position=position_dodge(width=0.9)) + theme_bw(base_size=12)

ggplot(backefRtry, aes(SDI, Ingroth, color=V1, fill=V1)) + geom_bar(stat="identity", position="dodge") + geom_errorbar(aes(ymin=Ingroth-se, ymax=Ingroth+se), width=0.4, position=position_dodge(width=0.9)) + theme_bw(base_size=12)

```

Warmer winter temperatures had a significant positive effect on both forest growth and ingrowth. Consequently, lower temperatures at higher altitudes negatively affected forest growth and ingrowth. 

Under more severe summer drought episodes, Scots pines were predicted to present lower growth and lower ingrowth when forest wood volume reaches higher levels, meaning competition increases. 




